---
title: "data simulation"
author: "Laura Symul"
date: "9/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
source("data_simulation_function.R")
```

# Simulation parameters

```{r seed-and-simulation-parameters}


N_ts <- 400 # number of time-series
L_ts <- 100 # length of each time-series
#sigma <- 0.5 # sd of normal noise on simulated input x
sigma <- 0.2
```


# Simulations

```{r data-simulation}

set.seed(1)

sim_data <- 
  map_dfr(
    .x = 1:N_ts,
    .f = simulate_data,
    L_ts = L_ts, 
    sigma = sigma
  )

```

```{r data-simulation-viz}

ggplot(sim_data %>% filter(ts_id < 10), 
       aes(x = t, y = var_value, col = y)) +
  geom_point(size = 0.2) +
  geom_line(size = 0.1) +
  facet_grid(y + ts_id ~ input_var) +
  guides(col = "none")

```

```{r dataset-balance}

ggplot(
  sim_data %>% 
    select(ts_id, y) %>% 
    distinct(),
  aes(x = y)
) +
  geom_bar()

```



# quick check that the output is not obvious from the average b and x

```{r check-mean-b-and-x}

ggplot(
  sim_data %>% 
    group_by(ts_id, input_var, y) %>% 
    summarize(mean_value = mean(var_value),
              .groups = "drop"), 
  aes(x = y, y = mean_value, fill = y)) +
  geom_violin(draw_quantiles = c(0.25,0.5,0.75)) +
  facet_grid(input_var ~ ., scales = "free") +
  guides(fill = "none")

```


```{r check-pca}

sim_data_wide <- 
  sim_data %>% 
  pivot_wider(
    names_from = c(input_var, t),
    values_from = var_value
  )

pca <- 
  prcomp(
    sim_data_wide %>% select(matches("_[0-9]*$"))
  )

plot(pca)

ggplot(
  pca$x %>%
    as_tibble() %>% 
    mutate(y = sim_data_wide$y),
  aes(x = PC1, y = PC2, col = y)
) +
  geom_point() +
  coord_fixed()

```

# export long and wide tables

```{r export}

write_csv(sim_data, file = "sim_data.csv")
write_csv(sim_data_wide, file = "sim_data_wide.csv")

```


# glm tests

```{r estimation of the parameters we know are relevant for output}

est_params <- 
  sim_data %>% 
  group_by(ts_id, y, input_var, b_prob, x_add_freq, x_offset) %>% 
  summarize(mean_value = mean(var_value),
            .groups = "drop") %>% 
  pivot_wider(
    names_from = input_var,
    values_from = mean_value,
  ) %>% 
  mutate(x = x-1)


get_freq <- 
  function(x){
    sp <- spectrum(x, plot = FALSE)
    sp$freq[sp$spec %>% which.max()]
  }


x_freq_hat <-
  sim_data %>% 
  filter(input_var == "x") %>% 
  mutate(x = var_value) %>% 
  select(ts_id, x) %>% 
  group_by(ts_id) %>% 
  summarize(
    x_freq = get_freq(x),
    .groups = "drop"
  ) 

est_params <- 
  est_params %>% 
  left_join(x_freq_hat, by = "ts_id") %>% 
  mutate(y_num = (y == 1)*1)

```



```{r}

lm(y_num ~ x_freq, # + b + x, 
   # family = "binomial", 
    data = est_params) %>% 
  summary()

```






# check that with knowledge of the model, we can estimate the output

Note: this output estimation needs to be adapted anytime the generative model for data simulation is modified.

```{r y-hat}

ggplot(est_output,
       aes(x = b_prob, y = b)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_point()

ggplot(est_output,
       aes(x = x_offset, y = x)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_point()

ggplot(est_output,
       aes(x = x_add_freq, y = x_freq)) +
  geom_abline(slope = 1, intercept = 0) +
  geom_point()

est_output <- 
  est_output %>% 
  mutate(
    y_hat = 
      # sign(b - 0.5) #*
      sign(0.5 - x_freq) # *
      # sign((b-0.5) * (0.5-x))
  ) %>% 
  mutate(
    y_hat = (y_hat > 0) * 1
  )

ggplot(est_output,
       aes(x = y, y = y_hat, fill = y)) +
  geom_violin()

table(est_output %>% select(y, y_hat))

```

It's not too bad. Would be better with a more accurate estimation of the x's frequency.






